---
title: "GTA: Greater Vancouver's Taste Analytics"
author: "Yin Hei Ernest Wong, Shiqi Zhang, Jiakun Li"
date: "`r Sys.Date()`"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Introduction

The **GTA (Greater Vancouver's Taste Analytics)** package is designed to help users analyze restaurant data in the Greater Vancouver Area using the **Yelp Fusion API**. This package provides functions to:
- Retrieve restaurant details (name, rating, address, cuisine type, etc.).
- Identify the highest-rated cuisines in a given area.
- Perform fuzzy matching on cuisine names.
- Visualize top restaurants based on ratings and reviews.

The **Yelp Fusion API** serves as the data source, allowing us to access up-to-date restaurant information.

---

# Installation

To install **GTA** from GitHub, use:

```{r}
library(remotes)
install_github("yhwong01/gtaR")
```

Once installed, load the package:
```{r}
library(gtaR)
```

---

Import Extra libraries:
```{r}
library(httr)        # For making API requests to Yelp
library(jsonlite)    # For parsing JSON responses
library(dplyr)       # For data wrangling
library(ggplot2)     # For data visualization
library(ggmap)       # For mapping restaurant locations
library(ggrepel)     # For avoiding label overlap in plots
library(stringdist)  # For fuzzy text matching

```

---

# Setting Up Your Yelp API Key

Before using any functions that access the Yelp API, you must set up your API key. Store it in your `.Renviron` file:

```{r}
Sys.setenv(YELP_API_KEY = "your_api_key")
```

Restart your R session for the changes to take effect.

---

## **1. search_restaurants()**

Queries Yelp for restaurants based on location and filters.

### **Required Parameters**

| Parameter       | Description                                                   | Default |
|----------------|---------------------------------------------------------------|---------|
| `location`     | The city or address for search.                               | "Vancouver, BC" |
| `term`         | Type of restaurant (e.g., "sushi", "Italian").                | "restaurant" |
| `limit`        | Number of results to return.                                  | `10` |
| `sort_by`      | Sorting criteria: "best_match", "rating", "review_count", "distance". | "rating" |
| `min_price`    | Minimum price level (`$` to `$$$$`).                          | `NULL` |
| `max_price`    | Maximum price level (`$` to `$$$$`).                          | `NULL` |
| `price_sort`   | Sort by price: "asc" or "desc".                           | `NULL` |

**Function Usage:**
```{r}
search_restaurants(location = "Vancouver, BC", term = "restaurant", limit = 10,  
sort_by = "rating", min_price = NULL, max_price = NULL, price_sort = NULL)
```

**Example:**
```{r}
restaurants <- search_restaurants(location = "Richmond, BC", term = "ramen", limit = 5)
print(restaurants)
```

---

# Functions and Usage

## **2. best_cuisine_in_area()**

Aggregates restaurant ratings to find the highest-rated cuisines in a specified location.

### **Required Parameters**

| Parameter  | Description                               | Default |
|-----------|-------------------------------------------|---------|
| `location` | The city or area for the search.        | "Vancouver, BC" |
| `limit`   | Number of restaurants to retrieve.      | `10` |

**Function Usage:**
```{r}
best_cuisine_in_area(location = "Vancouver, BC", limit = 10)
```

**Example:**
```{r}
best_cuisines <- best_cuisine_in_area("Burnaby, BC", limit = 15)
print(best_cuisines)
```

---

## **3. find_closest_cuisine_match()**

Uses fuzzy string matching to find the closest known cuisine type.

### **Required Parameters**

| Parameter    | Description                                        | Default |
|-------------|----------------------------------------------------|---------|
| `input_text` | A string representing the cuisine type to search. | None |

**Function Usage:**
```{r}
find_closest_cuisine_match(input_text)
```

**Example:**
```{r}
best_match <- find_closest_cuisine_match("Italien")  # Should return "Italian"
print(best_match)
```

---

## **4. get_restaurant_details()**

Retrieves details about a restaurant based on name or address.

### **Required Parameters**

| Parameter  | Description                                   | Default |
|-----------|-----------------------------------------------|---------|
| `query`   | Restaurant name or address.                  | None |
| `location` | The city or area for the search.            | "Vancouver, BC" |

**Function Usage:**
```{r}
get_restaurant_details(query, location = "Vancouver, BC")
```

**Example:**
```{r}
restaurant_info <- get_restaurant_details("Sushi Yama", "Vancouver, BC")
print(restaurant_info)
```

---


## **5. plot_best_restaurants()**

Visualizes the best restaurants based on ratings and review count.

### **Required Parameters**

| Parameter  | Description                           | Default |
|-----------|---------------------------------------|---------|
| `df`      | A tibble containing restaurant data. | None |
| `top_n`   | Number of top restaurants to display. | `10` |

**Function Usage:**
```{r}
plot_best_restaurants(df, top_n = 10)
```

**Example:**
```{r}
plot_best_restaurants(restaurants, top_n = 10)
```

**Example Output:**
```{r}
# A scatter plot of the top-rated restaurants
```

![plot_best_restaurants](https://github.com/user-attachments/assets/5e7f1d5c-95d5-4696-85de-d33ca03867a7)

---

## **6. list_available_locations()**

Returns a list of predefined cities available for restaurant searches.

### **Required Parameters**

| Parameter  | Description                        | Default |
|-----------|------------------------------------|---------|
| None      | Returns a predefined list of cities. | None |

**Function Usage:**
```{r}
list_available_locations()
```

**Example:**
```{r}
locations <- list_available_locations()
print(locations)
```

---

# Conclusion

The **GTA** package simplifies restaurant data retrieval and visualization for the **Greater Vancouver Area** using Yelp's API. Whether you want to find the best-rated cuisine, search for restaurants, or visualize trends, **GTA** provides an easy-to-use interface.

For more details, visit our [GitHub repository](https://github.com/yhwong01/gtaR).

---

# License

This package is licensed under the **MIT License**.

---